Short-term synaptic and neuronal plasticity act in synergy for deviance detection in spiking networks

### 2022-12-14 paper-1/manuscript.tex
Rev sec Results.STD (F7)

As noted above, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA-only condition examined in the previous section against the same network with both TA and STD. As shown in Figure \ref{fig6}, we found that the non-target response magnitude was reduced by the introduction of STD, as we would expect. Indeed, across networks, the response to all stimuli was reduced, regardless of context. Concomitantly, and in keeping with the above analysis, the levels of $V_{TA}$ before A trials also dropped.

\begin{figure*}%
    \centering
    \includegraphics{fig6.pdf}
    \caption{Contrast between the ablated model with only TA (shown in Figure \ref{fig5}) and the full model with both STD and TA. \textbf{a}, \textbf{b} Data shown in Figure \ref{fig5}ab, dev and msc columns, subtracted from the corresponding data for the full model. \textbf{c} Equivalent contrast (full - ablated model) in the response to A, in columns as above. \textbf{d} Network-wide mean $V_{TA}$ before A trials across networks. Left: Ablated and full model averages. Right: Difference between these model conditions. With the introduction of STD, $V_{TA}$ decreased more in dev than in msc (t = -3.49, p = 0.000342). \textbf{e} Mean number of spikes per target trial across networks. Left: Ablated and full model averages. Right: Difference between these model conditions. With the introduction of STD, the response decreased more in the std sequence than msc (t = -1.9, p = 0.0297), and more in msc than in dev (t = -2.99, p = 0.00168)}
    \label{fig6}
\end{figure*}

>>>>

% adding STD: What happens?
As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks with both TA and STD. On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown). However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence, this difference leading directly to the greater DD and SSA indices reported in Figure \ref{fig3}.

% Why does dXA increase?:
To understand the reason for this difference, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model. Firstly, as shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI). In most neurons, the contrast was greater in the full model, such that the difference (full - ablated) was positive, as indicated by the largely orange plot in the right column.
Following the logic established in the previous section, we may expect a greater target response in dev to be caused by lower $V_{TA}$ relative to msc. Indeed, comparing the contrast (dev - msc) in $V_{TA}$ before target trials across models (Figure \ref{fig7}b), we see that, while the contrast is negative (i.e., lower in dev) in both models, the magnitude of the contrast is greater in the full model. For clarity, the difference is therefore negative, as indicated by the largely blue plot in the right column.
Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response, which is exactly what we found: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.

This pattern is perhaps simpler to grasp in the aggregate across networks and stimuli. Firstly, the contrast (dev - msc) in the target response (Figure \ref{fig7}d, left) was positive in both models, but the full model contrast was greater than the ablated model contrast (t = 2.99, p = 0.00168; one-sided t-test, n = 120). Referring to the previous section, we locate the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342). Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921). In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc.

\begin{figure*}%
    \centering
    \includegraphics{fig7.pdf}
    \caption{}
    \label{fig7}
\end{figure*}
\textbf{Fig. 7} Adding STD to the model enhances the effect of TA.
\textbf{a} Contrast (dev - msc) of the response to A, in terms of mean number of spikes per trial, in the sample network, with each neuron shown in its spatial location. Left and middle column, full model with STD and TA, and ablated model with only TA, respectively. Right column, subtraction of these two. Stimulus locations are as shown in Figure \ref{fig5}a.
\textbf{b} Contrast of mean $V_{TA}$ at the start of target trials, in columns as in panel a.
\textbf{c} Contrast of the response to non-target stimuli, in columns as in panel a.
\textbf{d} Network-wide mean contrast of the response to target (left) and non-target stimuli (right) across networks and stimuli, arranged equivalently to the columns in panels a-c. Blue and green boxes represent the full and ablated model, respectively, and white boxes the difference between these two.
\textbf{e} Network-wide mean contrast of $V_{TA}$ before A trials across networks and stimuli, arranged as in panel d
